### 2024-06-14 paper/paper.tex
Initial commit

\PassOptionsToPackage{table}{xcolor}
\documentclass[review,onefignum,onetabnum,a4paper]{siamart190516}
\pdfoutput=1
\overfullrule=0pt

%------------------------------------------------------------------------------
% Header
%------------------------------------------------------------------------------
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath, amssymb, gensymb, physics, upgreek, siunitx, bbm}
\usepackage{tikz, pgfplots, graphicx, epstopdf, float, hyperref, xspace}
\usepackage[caption=false]{subfig}
\usepackage{enumitem, verbatim, listings}
\usepackage{mathtools, stmaryrd}
\usepackage{booktabs, array, ragged2e}
\usepackage[export]{adjustbox}
\usepackage{cleveref}

\sisetup{table-number-alignment=center, exponent-product=\times}
\newcommand{\crefrangeconjunction}{--}

\usetikzlibrary{trees, cd, babel}
\graphicspath{{./figures}}
\epstopdfsetup{outdir=./}

\lstset{basicstyle=\footnotesize\ttfamily}

\definecolor{paired1}{HTML}{a6cee3}
\definecolor{paired2}{HTML}{1f78b4}
\definecolor{paired3}{HTML}{b2df8a}
\definecolor{paired4}{HTML}{33a02c}
\definecolor{paired5}{HTML}{fb9a99}
\definecolor{paired6}{HTML}{e31a1c}
\definecolor{paired7}{HTML}{fdbf6f}
\definecolor{paired8}{HTML}{ff7f00}
\definecolor{paired9}{HTML}{cab2d6}
\definecolor{paired10}{HTML}{6a3d9a}

\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

\pgfplotsset{
  log x ticks with fixed point/.style={
      xticklabel={
        \pgfkeys{/pgf/fpu=true}
        \pgfmathparse{exp(\tick)}%
        \pgfmathprintnumber[fixed relative, precision=3]{\pgfmathresult}
        \pgfkeys{/pgf/fpu=false}
      }
  },
  log y ticks with fixed point/.style={
      yticklabel={
        \pgfkeys{/pgf/fpu=true}
        \pgfmathparse{exp(\tick)}%
        \pgfmathprintnumber[fixed relative, precision=3]{\pgfmathresult}
        \pgfkeys{/pgf/fpu=false}
      }
  }
}

% Classic symbols
\newcommand{\reals}{\mathbb{R}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\dx}{\,\d\mathbf{x}}
\newcommand{\md}{\,\d}
\newcommand{\bigo}[1]{\mathcal{O}(#1)}
\DeclareMathOperator*{\argmin}{arg\,min}

\let\grad\undefined
\let\curl\undefined
\let\div\undefined
\let\nullop\undefined
\let\d\undefined
\let\tr\undefined
\DeclareMathOperator{\grad}{grad}
\DeclareMathOperator{\curl}{curl}
\DeclareMathOperator{\div}{div}
\DeclareMathOperator{\nullop}{null}
\DeclareMathOperator{\d}{d}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\Exists}{\exists}
\DeclareMathOperator{\Forall}{\forall}

\DeclareMathOperator{\refgrad}{\widehat{\grad}}
\DeclareMathOperator{\refcurl}{\widehat{\curl}}
\DeclareMathOperator{\refdiv}{\widehat{\div}}

\newcommand{\Hgrad}{H(\grad)}
\newcommand{\Hcurl}{H(\curl)}
\newcommand{\Hdiv}{H(\div)}
\newcommand{\Ltwo}{L^2}

\newcommand{\mesh}{\mathcal{T}_h}
\newcommand{\skeleton}{\mathcal{E}_h}
\newcommand{\refline}{\hat{\mathcal{I}}}
\newcommand{\jac}{\mathrm{D}F}
\newcommand{\pullback}{\mathcal{F}}
\newcommand{\betah}{\tilde{\beta}}

% Vector notation
\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\uvec}[1]{\mathbf{\hat{#1}}}
\newcommand{\conj}[1]{{\overline{#1}}}
\DeclareMathOperator{\blockdiag}{blockdiag}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\spn}{span}
\DeclareMathOperator{\vecop}{vec}
\newcommand{\eye}{{\mathbb{I}}}
\newcommand{\kron}{\otimes}
\newcommand{\jump}[1]{\left\llbracket #1 \right\rrbracket}
\newcommand{\avg}[1]{\left\{ #1 \right\}}
\newcommand{\broken}[1]{\tilde{#1}}

\newcommand{\be}{\vec{e}}
\newcommand{\bu}{\vec{u}}
\newcommand{\bn}{\vec{n}}
\newcommand{\bm}{\vec{m}}
\newcommand{\bv}{\vec{v}}
\renewcommand{\bf}{\vec{f}}
\newcommand{\bq}{\vec{q}}
\newcommand{\br}{\vec{r}}
\newcommand{\bx}{\vec{x}}
\newcommand{\bw}{\vec{w}}
\newcommand{\by}{\vec{y}}
\newcommand{\bz}{\vec{z}}
\newcommand{\bB}{\vec{B}}
\newcommand{\bC}{\vec{C}}
\newcommand{\bE}{\vec{E}}
\newcommand{\bF}{\vec{F}}
\newcommand{\bJ}{\vec{J}}
\newcommand{\bP}{\vec{P}}
\newcommand{\bT}{\vec{T}}
\newcommand{\bX}{\vec{X}}

\newcommand{\ub}{\underline{b}}
\newcommand{\uc}{\underline{c}}
\newcommand{\ue}{\underline{e}}
\newcommand{\uf}{\underline{f}}
\newcommand{\ug}{\underline{g}}
\newcommand{\up}{\underline{p}}
\newcommand{\uq}{\underline{q}}
\newcommand{\ur}{\underline{r}}
\newcommand{\uu}{\underline{u}}
\newcommand{\uv}{\underline{v}}
\newcommand{\uw}{\underline{w}}
\newcommand{\ux}{\underline{x}}
\newcommand{\uy}{\underline{y}}
\newcommand{\uz}{\underline{z}}
\newcommand{\ut}{\underline{t}}
\newcommand{\ulambda}{\underline{\lambda}}

\newcommand{\uB}{\underline{B}}
\newcommand{\ubb}{\underline{\vec{b}}}
\newcommand{\ubq}{\underline{\vec{q}}}
\newcommand{\ubu}{\underline{\vec{u}}}
\newcommand{\ubv}{\underline{\vec{v}}}
\newcommand{\ubf}{\underline{\vec{f}}}
\newcommand{\ubg}{\underline{\vec{g}}}

\newcommand{\xhat}{\hat{x}}
\newcommand{\rhat}{\hat{r}}
\newcommand{\shat}{\hat{s}}
\newcommand{\Khat}{\hat{K}}
\newcommand{\Bhat}{\hat{B}}
\newcommand{\Ahat}{\hat{A}}
\newcommand{\Dhat}{\hat{D}}
\newcommand{\Shat}{\hat{S}}
\newcommand{\Ghat}{\hat{G}}

\renewcommand{\P}{\mathrm{P}}
\newcommand{\DP}{\mathrm{DP}}
\newcommand{\Q}{\mathrm{Q}}
\newcommand{\DQ}{\mathrm{DQ}}
\newcommand{\RTCF}{\mathrm{RTCF}}
\newcommand{\RTCE}{\mathrm{RTCE}}
\newcommand{\NCF}{\mathrm{NCF}}
\newcommand{\NCE}{\mathrm{NCE}}

%\newtheorem{theorem}{Theorem}
\newtheorem{remark}{Remark}

%------------------------------------------------------------------------------
% Document
%------------------------------------------------------------------------------

\title{
   Fast solvers for the high-order FEM simplicial de Rham complex
\thanks{Submitted to the editors XX/YY/2024.
      \funding{
         PDB and PEF were supported by EPSRC grant EP/W026260/1.
      }
   }
}

\author{Pablo D.\ Brubeck\thanks{
Mathematical Institute,
University of Oxford,
Oxford, UK (\email{brubeckmarti@maths.ox.ac.uk})
%\orcid{0000-0002-3824-0080}
}
\and Patrick E.\ Farrell\thanks{
Mathematical Institute,
University of Oxford,
Oxford UK (\email{patrick.farrell@maths.ox.ac.uk})
%\orcid{0000-0002-1241-7060}
}
}

\headers{Patch solvers for simplicial $p$-FEM}{P.~D.~Brubeck and P.~E.~Farrell and R.~C.~Kirby}

\begin{document}

\numberwithin{equation}{section}
\maketitle

\begin{abstract}
This is the abstract.
\end{abstract}

\begin{keywords}
   preconditioning, de Rham complex, simplex, high-order, additive Schwarz
\end{keywords}

\begin{AMS}
   65F08, 65N35, 65N55
\end{AMS}

%\begin{DOI}
%10.1137/21XXXXXXX
%\end{DOI}

\section{Introduction} \label{sec:introduction}

\section{Conclusion} \label{sec:conclusion}

\bibliographystyle{siamplain}
\bibliography{references}
\end{document}

### 2024-06-14 paper/paper.tex
Define degrees of freedom and space decompositions

@@
 \newcommand{\bn}{\vec{n}}
+\newcommand{\bt}{\vec{t}}
 \newcommand{\bm}{\vec{m}}
@@
 \newcommand{\DP}{\mathrm{DP}}
-\newcommand{\Q}{\mathrm{Q}}
-\newcommand{\DQ}{\mathrm{DQ}}
-\newcommand{\RTCF}{\mathrm{RTCF}}
-\newcommand{\RTCE}{\mathrm{RTCE}}
-\newcommand{\NCF}{\mathrm{NCF}}
-\newcommand{\NCE}{\mathrm{NCE}}
+\newcommand{\CG}{\mathrm{CG}}
+\newcommand{\DG}{\mathrm{DG}}
+\newcommand{\Ned}{\mathrm{Ned}^{1}}
+\newcommand{\RT}{\mathrm{RT}}
+\newcommand{\BDM}{\mathrm{BDM}}
+\newcommand{\NedTwo}{\mathrm{Ned}^{2}}
+
+\newcommand{\Nedelec}{Ned\'el\'ec~}
 
@@
    Fast solvers for the high-order FEM simplicial de Rham complex
-\thanks{Submitted to the editors XX/YY/2024.
+\thanks{Submitted to the editors July XX, 2024.
       \funding{
@@
 University of Oxford,
-Oxford UK (\email{patrick.farrell@maths.ox.ac.uk})
+Oxford, UK (\email{patrick.farrell@maths.ox.ac.uk})
+%\orcid{0000-0002-1241-7060}
+}
+\and Robert C.\ Kirby\thanks{
+Department of Mathematics,
+Baylor University,
+Waco, TX, USA (\email{robert\_kirby@baylor.edu})
 %\orcid{0000-0002-1241-7060}
@@
 \begin{abstract}
-This is the abstract.
+We present new high-order finite elements discretizing the $L^2$ de Rham
+complex on triangular and tetrahedral meshes. The finite elements discretize
+the same spaces as usual, but with different basis functions. They allow for
+fast linear solvers based on static condensation and space decomposition
+methods. The new elements build upon the definition of degrees of freedom
+given in L.~Demkowicz, P.~Monk, L.~Vardapetyan, and W.~Rachowicz.~
+\textit{De Rham diagram for $hp$ finite element spaces}. Comput.~Math.~Appl.,
+39(7-8):29--38, 2000., and consist of integral moments
+on an equilateral reference simplex with respect to a numerically computed
+polynomial basis that is orthogonal in both the $L^2$- and
+$H(\mathrm{d})$-inner products ($\mathrm{d} \in \{\mathrm{grad},
+\mathrm{curl}, \mathrm{div}\}$). On the reference equilateral simplex, the
+resulting stiffness matrix has diagonal interior block, and does not couple
+together the interior and interface degrees of freedom. Thus, on the
+reference simplex, the Schur complement resulting from elimination of
+interior degrees of freedom is simply the interface block itself. This
+sparsity is not preserved on arbitrary cells mapped from the reference cell.
+Nevertheless, the interior-interface coupling is weak because it is only
+induced by the geometric transformation. We devise a preconditioning
+strategy by neglecting the interior-interface coupling. We precondition the
+interface Schur complement with the interface block, and simply apply
+point-Jacobi to precondition the interior block. The combination of this
+approach with a space decomposition method on small subdomains constructed
+around vertices, edges, and faces allows us to efficiently solve the
+canonical Riesz maps in $H(\mathrm{grad})$, $H(\mathrm{curl})$, and
+$H(\mathrm{div})$, at very high order. We empirically demonstrate iteration
+counts that are robust with respect to the polynomial degree. 
 \end{abstract}
@@
 
+\section{Sparsity-promoting discretizations} \label{sec:dofs}
+
+We discretize $\Ltwo$-de Rham complex in triangular and tetrahedral meshes.
+
+\begin{figure}[htbp] 
+\centering
+\begin{tikzcd}
+  \Hgrad \arrow[r, "\grad"] \arrow[d]  & \Hcurl \arrow[r, "\curl"]
+  \arrow[d] & \Hdiv \arrow[r, "\div"]  \arrow[d] & \Ltwo  \arrow[d] \\
+  \CG_p \arrow[r, "\grad"] & \Ned_p \arrow[r, "\curl"] & 
+   \RT_p \arrow[r, "\div"] & \DG_{p-1}
+\end{tikzcd}
+
+\begin{tikzcd}
+  \Hgrad \arrow[r, "\grad"] \arrow[d]  & \Hcurl \arrow[r, "\curl"]
+  \arrow[d] & \Hdiv \arrow[r, "\div"]  \arrow[d] & \Ltwo  \arrow[d] \\
+   \CG_p \arrow[r, "\grad"] & \NedTwo_{p-1} \arrow[r, "\curl"] & 
+   \BDM_{p-2} \arrow[r, "\div"] & \DG_{p-3}
+\end{tikzcd}
+\caption{The $\Ltwo$-de Rham complex, and the finite element subcomplexes of the first and second kind.}
+\end{figure}
+
+We follow the definition of interpolation degrees of freedom from \cite{demkowicz00}
+to construct dual bases that promote orthgonality in the $\Ltwo(\Khat)$ and $H(\mathrm{d}, \Khat)$ inner
+products.
+
+\subsection{Sparsity-promoting basis for $\Hgrad$}
+
+Let $\Delta^d$ be the equilateral $d$-simplex.
+We define a basis for the dual of $\CG_p$ as 
+point evaluations at $V \in \text{vertices}(\Delta^d)$,
+\begin{equation}
+   \ell^V(v) = v(\bx_V),
+\end{equation}
+and for each sub-entity $S \in \text{edges}(\Delta^d) \cup \text{faces}(\Delta^d) \cup \text{interior}(\Delta^d)$
+we take integral moments of surface gradients, 
+\begin{equation}
+   \ell^S_j(v) = (\grad_S\phi^S_j, \grad_S v)_{S},
+\end{equation}
+where $\grad_S$ is the tangential gradient on $S$,
+and $\{\phi^S_j\}$ is a basis for
+$\P_{p,0}(S) = \{ v \in \P_p(S): v = 0 \text{ on } \partial S \}$, such that
+\begin{equation} \label{eq:hgrad-fdm}
+   (\grad_S\phi^S_j, \grad_S\phi^S_i)_{S} = \delta_{ij}, \quad
+   (\phi^S_j, \phi^S_i)_{S} = \lambda_j\delta_{ij}.
+\end{equation}
+The eigenbases $\{\phi^S_j\}$ are numerically computed offline and only once 
+on the reference interval, triangle, and tetrahedron.
+
+\subsection{Sparsity-promoting basis for $\Hcurl$}
+
+Recall the definition of the $\Hcurl$-confroming finite element spaces
+\begin{equation}
+V^1_{p, p_E}(K) = \{v\in [\P_p(K)]^d : v\cdot\bt \in \P_{p_E}(E) \}
+\end{equation}
+for $p_E=p-1$ we obtain the \Nedelec space of the first kind $\Ned_p(K)$, while
+$p_E=p$ gives the \Nedelec space of the second kind $\NedTwo_p(K) = [\P_p(K)]^d$. 
+These spaces enforce continuity of the tangential trace across mesh interfaces.
+
+We define a basis for the dual of $V^1_{p, p_E}(K)$ as 
+tangential moments along $E\in \text{edges}(\Delta^d)$,
+\begin{equation}
+   \ell^E_j(v) = (q_j, v\cdot \bt)_E, \quad q_j \in \P_{p_E}(E), 
+\end{equation}
+and for each sub-entity $S \in \text{faces}(\Delta^d) \cup \text{interior}(\Delta^d)$,
+\begin{align}
+   \ell^{S,0}_j(v) &= (\grad_S\phi^S_j, v)_{S}, \\
+   \ell^{S,1}_j(v) &= (\curl_S\Phi^S_j, \curl_S v)_{S},
+\end{align}
+where $\curl_S$ is the tangential curl on $S$,
+$\{\phi^S_j\}$ are the $H_0(\grad, S)$ bases constructed in \eqref{eq:hgrad-fdm},
+and $\{\curl_S\Phi^S_j\}$ is a basis for $\curl_S \mathbb{X}_{p_S}(S)$,
+$\mathbb{X}_{p_S}(S) = [\P_{p_S}(S)]^d \cap H_0(\curl, S)$, such that
+\begin{equation} \label{eq:hcurl-fdm}
+   (\curl_S\Phi^S_j, \curl_S\Phi^S_i)_{S} = \delta_{ij}, \quad
+   (\Phi^S_j, \Phi^S_i)_{S} = \lambda_j\delta_{ij}, \quad
+   \Phi^S_j\times\bn = 0 \text{ on }\partial S.
+\end{equation}
+
+\subsection{Sparsity-promoting basis for $\Hdiv$}
+
+Recall the definition of the $\Hdiv$-confroming finite element spaces
+\begin{equation}
+V^2_{p, p_F}(K) = \{v\in [\P_p(K)]^d : v\cdot\bn \in \P_{p_F}(F) \}
+\end{equation}
+for $p_F=p-1$ we obtain the Raviart--Thomas space $\RT_p(K)$, while
+$p_F=p$ gives the Brezzi--Douglas--Marini space $\BDM_p(K) = [\P_p(K)]^d$. 
+These spaces enforce continuity of the normal trace across mesh interfaces.
+
+We define a basis for the dual of $V^2_{p, p_F}(K)$ as 
+normal moments on $F\in \text{faces}(\Delta^d)$,
+\begin{equation}
+   \ell^F_j(v) = (q_j, v\cdot \bn)_F, \quad q_j \in \P_{p_F}(F),
+\end{equation}
+and on the $\text{interior}(\Delta^d) = K$: 
+\begin{align}
+   \ell^{K,0}_j(v) &= (\curl\Phi^K_j, v)_{K}, \\
+   \ell^{K,1}_j(v) &= (\div\Psi^K_j, \div v)_{K},
+\end{align}
+where $\{\Phi^K_j\}$ are the $H_0(\curl, S)$ bases constructed in \eqref{eq:hcurl-fdm},
+and $\{\div\Psi^K_j\}$ is a basis for $\div \mathbb{Y}_p(K)$,
+$\mathbb{Y}_{p}(K) = [\mathbb{P}_p(K)]^d \cap H_0(\div,K)$, such that
+\begin{equation} \label{eq:hdiv-fdm}
+   (\div\Psi^K_j, \div\Psi^K_i)_{K} = \delta_{ij}, \quad
+   (\Psi^K_j, \Psi^K_i)_{K} = \lambda_j\delta_{ij}, \quad
+   \Psi^K_j\cdot\bn = 0 \text{ on }\partial K.
+\end{equation}
+
+\section{Space decompositions}
+
+We follow 
+the low-order space decompositions studied in \cite{arnold00}
+and extend them to the high-order with static-condensation.
+
+For 0-forms we construct patches on vertex-stars,
+\begin{equation}
+   V^0_{h,p} = \left.V^0_{h,p}\right|_\mathcal{I} + V^0_{h,1} 
+   + \sum_{v\in E^0(\mesh)} \left.\tilde{V}^0_{h,p}\right|_{\star v} 
+\end{equation}
+
+For 1-forms we construct edge-star patches
+plus the gradient of vertex-star patches on 0-forms
+\begin{equation}
+   V^1_{h,p} = \left.V^1_{h,p}\right|_\mathcal{I} +  V^1_{h,1}
+   + \sum_{e\in E^1(\mesh)} \left.\tilde{V}^1_{h,p}\right|_{\star e} 
+   + \sum_{v\in E^0(\mesh)} \mathrm{d} \left.\tilde{V}^0_{h,p}\right|_{\star v} 
+\end{equation}
+
+For 2-forms we construct edge-star patches,
+\begin{equation}
+   V^2_{h,p} = \left.V^2_{h,p}\right|_\mathcal{I} +  V^2_{h,1} 
+   + \sum_{e\in E^1(\mesh)} \left.\tilde{V}^2_{h,p}\right|_{\star e}. 
+\end{equation}
+
+\section{Numerical results} \label{sec:results}
+
+\subsection{Unit cube tests}
+Here we want to address questions about $h$ and $p$ robustness of the space
+decomposition. Also how does this method compare with other standard bases in
+terms of not only of runtime, memory, and FLOPs, but also conditioning, and
+approximation error.
+
+\subsection{Mixed Poisson}
+Is there a substantial advantage in the first order reformulation of Poisson solved
+with a block Riesz map preconditioner over the $C^0$ formulation?
+Since the normal trace of the $\Hdiv$ space is just a scalar, we expect that the
+statically-condensed edge patches should couple fewer cells.
+
+\subsection{Complicated geometries and adaptive mesh refinement}
+How does the solver perform on complicated and adaptively-refined meshes?
+
 \section{Conclusion} \label{sec:conclusion}

### 2024-06-14 paper/paper.tex
Notation for bubbles

@@
 
-We follow the definition of interpolation degrees of freedom from \cite{demkowicz00}
-to construct dual bases that promote orthgonality in the $\Ltwo(\Khat)$ and $H(\mathrm{d}, \Khat)$ inner
-products.
+We follow the definition of interpolation degrees of freedom from
+\cite{demkowicz00} to construct dual bases that promote orthogonality in the
+$\Ltwo(\Khat)$ and $H(\mathrm{d}, \Khat)$ inner products.
+
+We denote the space of polynomials with vanishing boundary trace as
+$\mathbb{B}_{p}(\mathrm{d}, K) = [\P_{p}(K)]^d \cap H_0(\mathrm{d}, K)$.
 
@@
 where $\grad_S$ is the tangential gradient on $S$,
-and $\{\phi^S_j\}$ is a basis for
-$\P_{p,0}(S) = \{ v \in \P_p(S): v = 0 \text{ on } \partial S \}$, such that
+and $\{\phi^S_j\}$ is a basis for $\mathbb{B}_{p}(\grad, S)$, such that
 \begin{equation} \label{eq:hgrad-fdm}
@@
 $\{\phi^S_j\}$ are the $H_0(\grad, S)$ bases constructed in \eqref{eq:hgrad-fdm},
-and $\{\curl_S\Phi^S_j\}$ is a basis for $\curl_S \mathbb{X}_{p_S}(S)$,
-$\mathbb{X}_{p_S}(S) = [\P_{p_S}(S)]^d \cap H_0(\curl, S)$, such that
+and $\{\curl_S\Phi^S_j\}$ is a basis for $\curl_S \mathbb{B}_{p_S}(\curl, S)$
+such that
 \begin{equation} \label{eq:hcurl-fdm}
@@
 
-We define a basis for the dual of $V^2_{p, p_F}(K)$ as 
-normal moments on $F\in \text{faces}(\Delta^d)$,
+We define a basis for the dual of $V^2_{p, p_F}(K)$ as normal moments on $F\in
+\text{faces}(\Delta^d)$,
 \begin{equation}
@@
 \end{align}
-where $\{\Phi^K_j\}$ are the $H_0(\curl, S)$ bases constructed in \eqref{eq:hcurl-fdm},
-and $\{\div\Psi^K_j\}$ is a basis for $\div \mathbb{Y}_p(K)$,
-$\mathbb{Y}_{p}(K) = [\mathbb{P}_p(K)]^d \cap H_0(\div,K)$, such that
+where $\{\Phi^K_j\}$ are the $H_0(\curl, S)$ bases constructed in
+\eqref{eq:hcurl-fdm}, and $\{\div\Psi^K_j\}$ is a basis for $\div
+\mathbb{B}_p(\div, K)$ such that
 \begin{equation} \label{eq:hdiv-fdm}
-   (\div\Psi^K_j, \div\Psi^K_i)_{K} = \delta_{ij}, \quad
+   (\div\Psi^K_j, \div\Psi^K_i)_K = \delta_{ij}, \quad
    (\Psi^K_j, \Psi^K_i)_{K} = \lambda_j\delta_{ij}, \quad

### 2024-06-14 paper/paper.tex
Minor edits

@@
 methods. The new elements build upon the definition of degrees of freedom
-given in L.~Demkowicz, P.~Monk, L.~Vardapetyan, and W.~Rachowicz.~
-\textit{De Rham diagram for $hp$ finite element spaces}. Comput.~Math.~Appl.,
-39(7-8):29--38, 2000., and consist of integral moments
+given in [L.~Demkowicz, P.~Monk, L.~Vardapetyan, and W.~Rachowicz,
+Comput.~Math.~Appl.,
+39(7-8) (2000), pp.~29--38], and consist of integral moments
 on an equilateral reference simplex with respect to a numerically computed
@@
 $H(\mathrm{d})$-inner products ($\mathrm{d} \in \{\mathrm{grad},
-\mathrm{curl}, \mathrm{div}\}$). On the reference equilateral simplex, the
+\mathrm{curl}, \mathrm{div}\}$). 
+On the reference equilateral simplex, the
 resulting stiffness matrix has diagonal interior block, and does not couple
@@
 the low-order space decompositions studied in \cite{arnold00}
-and extend them to the high-order with static-condensation.
+and extend them to high-order with static-condensation.
 

### 2024-06-14 paper/paper.tex
Remark on avoding  the FDM eigenproblem for the second type interior DOFs at the end of the complex

--- 2024-06-14 paper/paper.tex
@@
 approach with a space decomposition method on small subdomains constructed
-around vertices, edges, and faces allows us to efficiently solve the
+around vertices and edges allows us to efficiently solve the
 canonical Riesz maps in $H(\mathrm{grad})$, $H(\mathrm{curl})$, and
@@
 
-We discretize $\Ltwo$-de Rham complex in triangular and tetrahedral meshes.
+We discretize $\Ltwo$-de Rham complex on triangular and tetrahedral meshes.
 
@@
 
-Recall the definition of the $\Hcurl$-confroming finite element spaces
+Recall the definition of the $\Hcurl$-conforming finite element spaces
 \begin{equation}
@@
    \ell^{S,0}_j(v) &= (\grad_S\phi^S_j, v)_{S}, \\
-   \ell^{S,1}_j(v) &= (\curl_S\Phi^S_j, \curl_S v)_{S},
+   \ell^{S,1}_j(v) &= (\curl_S\Phi^S_j, \curl_S v)_{S}, \label{eq:hcurl-interior-dofs}
 \end{align}
@@
 \end{equation}
+On a 2D face $F$ (or the interior of a 2D cell),
+it is not necessary to solve for $\{\Phi^F_j\}$, as the degrees of freedom
+\eqref{eq:hcurl-interior-dofs} only involve the scalar-valued 
+$\curl_F\Phi^F_j$ which spans $\P_{p_F-1}(F)\setminus\P_0(F)$,
+hence the standard $\Ltwo$-orthogonal Dubiner basis may be used in place of $\curl_F\Phi^F_j$.
 
@@
 
-Recall the definition of the $\Hdiv$-confroming finite element spaces
+Recall the definition of the $\Hdiv$-conforming finite element spaces
 \begin{equation}
@@
    \ell^{K,0}_j(v) &= (\curl\Phi^K_j, v)_{K}, \\
-   \ell^{K,1}_j(v) &= (\div\Psi^K_j, \div v)_{K},
+   \ell^{K,1}_j(v) &= (\div\Psi^K_j, \div v)_{K}, \label{eq:hdiv-interior-dofs}
 \end{align}
@@
 \end{equation}
+It is not necessary to solve for $\{\Psi^K_j\}$, as the degrees of freedom
+\eqref{eq:hdiv-interior-dofs} only involve $\div\Psi^K_j$, 
+which spans $\P_{p-1}(K)\setminus\P_0(K)$. Hence the standard
+$\Ltwo$-orthogonal Dubiner basis may be used in place of $\div\Psi^K_j$.
+
+\subsection{Generally-mapped elements}
+
+The dual bases described above yield a sparse basis for $V^k_p(K)$ only
+if $K$ is equilateral.
 

### 2024-06-14 paper/paper.tex
remove remark

@@
 \end{equation}
-On a 2D face $F$ (or the interior of a 2D cell),
-it is not necessary to solve for $\{\Phi^F_j\}$, as the degrees of freedom
-\eqref{eq:hcurl-interior-dofs} only involve the scalar-valued 
-$\curl_F\Phi^F_j$ which spans $\P_{p_F-1}(F)\setminus\P_0(F)$,
-hence the standard $\Ltwo$-orthogonal Dubiner basis may be used in place of $\curl_F\Phi^F_j$.
 
@@
 \end{equation}
-It is not necessary to solve for $\{\Psi^K_j\}$, as the degrees of freedom
-\eqref{eq:hdiv-interior-dofs} only involve $\div\Psi^K_j$, 
-which spans $\P_{p-1}(K)\setminus\P_0(K)$. Hence the standard
-$\Ltwo$-orthogonal Dubiner basis may be used in place of $\div\Psi^K_j$.
 
@@
 
-The dual bases described above yield a sparse basis for $V^k_p(K)$ only
-if $K$ is equilateral.
+On an equilateral simplex, the interior block of the mass and stiffness matrices
+is fully diagonal, moreover the interior-interface blocks in the stiffness 
+vanish for generally mapped elements.
+
+The dual bases described above yield interior-orthogonal primal bases for $V^k_p(K)$ only
+if $K$ is equilateral. For generally mapped triangles, the spectral equivalence constant of
+a space decomposition that decouples the interior degrees of freedom should depend on how
+far the element departs from being equilateral, i.e., the shape regularity parameter.
 

### 2024-06-17 paper/paper.tex
...

@@
 \usepackage{cleveref}
+\usepackage{todonotes}
 
@@
 
+\newcommand{\pef}[1]{\todo[inline,color=blue!40]{PEF: \#1}}
+
 %\newtheorem{theorem}{Theorem}
@@
 $H(\mathrm{div})$, at very high order. We empirically demonstrate iteration
-counts that are robust with respect to the polynomial degree. 
+counts that are robust with respect to the polynomial degree, with a computational complexity of $\mathcal{O}(p^{6})$ flops in three dimensions
+instead of the na\"ive $\mathcal{O}(p^{9})$.
 \end{abstract}
@@
 
-We discretize $\Ltwo$-de Rham complex on triangular and tetrahedral meshes.
+We discretize the $\Ltwo$ de Rham complex on triangular and tetrahedral meshes with finite element subcomplexes of the first and second kinds.
 
@@
 \begin{tikzcd}
-  \Hgrad \arrow[r, "\grad"] \arrow[d]  & \Hcurl \arrow[r, "\curl"]
-  \arrow[d] & \Hdiv \arrow[r, "\div"]  \arrow[d] & \Ltwo  \arrow[d] \\
   \CG_p \arrow[r, "\grad"] & \Ned_p \arrow[r, "\curl"] & 
-   \RT_p \arrow[r, "\div"] & \DG_{p-1}
-\end{tikzcd}
-
-\begin{tikzcd}
-  \Hgrad \arrow[r, "\grad"] \arrow[d]  & \Hcurl \arrow[r, "\curl"]
-  \arrow[d] & \Hdiv \arrow[r, "\div"]  \arrow[d] & \Ltwo  \arrow[d] \\
+   \RT_p \arrow[r, "\div"] & \DG_{p-1} \\
+  \Hgrad \arrow[r, "\grad"] \arrow[u] \arrow[d]  & \Hcurl \arrow[r, "\curl"]
+  \arrow[d] \arrow[u] & \Hdiv \arrow[r, "\div"] \arrow[u] \arrow[d] & \Ltwo \arrow[u] \arrow[d] \\
    \CG_p \arrow[r, "\grad"] & \NedTwo_{p-1} \arrow[r, "\curl"] & 
@@
 \end{tikzcd}
-\caption{The $\Ltwo$-de Rham complex, and the finite element subcomplexes of the first and second kind.}
+\caption{The $\Ltwo$-de Rham complex (middle), and the finite element subcomplexes of the first (above) and second kinds (below).}
 \end{figure}

### 2024-06-17 paper/paper.tex
Some work on section 2

@@
 \DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\union}{\cup}
 
@@
 
-\newcommand{\pef}[1]{\todo[inline,color=blue!40]{PEF: \#1}}
+\newcommand{\pef}[1]{\todo[inline,color=blue!40]{PEF: #1}}
 
@@
 \end{tikzcd}
-\caption{The $\Ltwo$-de Rham complex (middle), and the finite element subcomplexes of the first (above) and second kinds (below).}
+\caption{The $\Ltwo$ de Rham complex (middle), and the finite element subcomplexes of the first (above) and second kinds (below).}
 \end{figure}
+\pef{Fix vertical alignment}
+
+For each finite element subcomplex built on a mesh $\mesh$ of a domain $\Omega \subset \mathbb{R}^d$, we denote the spaces discretizing $k$-forms $H\Lambda^k(\Omega)$ with polynomial degree $p$ as $X^k_p(\mesh)$.
+These are built from cellwise spaces via the standard pullback construction from a reference cell $\Khat$:
+\begin{equation}
+X^k_p(\mesh) = \{ v \in H\Lambda^k(\Omega): \Forall K \in \mesh \Exists \hat{v} \in X^k_p(\Khat) \text{ s.t.} \left.v\right|_K = \pullback^k_K(\hat{v}) \},
+\end{equation}
+where $\pullback^k_K$ denotes the appropriate pullback that preserves continuity of the traces of a $k$-form across cell facets.
 
 We follow the definition of interpolation degrees of freedom from
-\cite{demkowicz00} to construct dual bases that promote orthogonality in the
+Demkowicz et al.~\cite{demkowicz00} to construct dual bases that promote orthogonality in the
 $\Ltwo(\Khat)$ and $H(\mathrm{d}, \Khat)$ inner products.
+In the construction of the degrees of freedom we will employ bubble subspaces on a given mesh entity $S \subset \Khat$ with zero trace, defined as
+\begin{equation}
+B^k_p(S) \coloneqq \{v \in X^k_p(S): \tr v = 0 \text{ on } \partial S\}.
+\end{equation}
 
-We denote the space of polynomials with vanishing boundary trace as
-$\mathbb{B}_{p}(\mathrm{d}, K) = [\P_{p}(K)]^d \cap H_0(\mathrm{d}, K)$.
+Throughout we choose $\Khat$ to be an equilateral simplex, as this ...
 
@@
 
-Let $\Delta^d$ be the equilateral $d$-simplex.
-We define a basis for the dual of $\CG_p$ as 
-point evaluations at $V \in \text{vertices}(\Delta^d)$,
+On $\Khat$ we discretize $\Hgrad$ with the space
 \begin{equation}
-   \ell^V(v) = v(\bx_V),
+X^0_{p}(\Khat) = \P_p(\Khat),
 \end{equation}
-and for each sub-entity $S \in \text{edges}(\Delta^d) \cup \text{faces}(\Delta^d) \cup \text{interior}(\Delta^d)$
+where $\P_p(\Khat)$ is the space of polynomials of total degree less than or equal to $p$.
+We define a basis for the dual of $X^0_p(\Khat)$ as
+point evaluations at $V \in \text{vertices}(\Khat)$,
+\begin{equation}
+   \ell^V(v) = v(V),
+\end{equation}
+and for each sub-entity $S \in \text{edges}(\Khat) \union \text{faces}(\Khat) \union \text{interior}(\Khat)$
 we take integral moments of surface gradients, 
@@
 where $\grad_S$ is the tangential gradient on $S$,
-and $\{\phi^S_j\}$ is a basis for $\mathbb{B}_{p}(\grad, S)$, such that
+and $\{\phi^S_j\}$ is a basis for $B^0_p(S)$
+such that
 \begin{equation} \label{eq:hgrad-fdm}
@@
 We define a basis for the dual of $V^1_{p, p_E}(K)$ as 
-tangential moments along $E\in \text{edges}(\Delta^d)$,
+tangential moments along $E\in \text{edges}(\Khat)$,
 \begin{equation}
@@
 \end{equation}
-and for each sub-entity $S \in \text{faces}(\Delta^d) \cup \text{interior}(\Delta^d)$,
+and for each sub-entity $S \in \text{faces}(\Khat) \cup \text{interior}(\Khat)$,
 \begin{align}
@@
 We define a basis for the dual of $V^2_{p, p_F}(K)$ as normal moments on $F\in
-\text{faces}(\Delta^d)$,
+\text{faces}(\Khat)$,
 \begin{equation}
@@
 \end{equation}
-and on the $\text{interior}(\Delta^d) = K$: 
+and on the $\text{interior}(\Khat) = K$: 
 \begin{align}

### 2024-06-18 paper/paper.tex
Sketch for the introduction

@@
 
+Main challenges, motivations
+\begin{itemize}
+\item In high-order FEM, the curse of dimensionality affects operator
+   evaluation and relaxation methods in multigrid and domain-decomposition.  
+\item For tensor product cells,
+   a tensor product basis will expose the structure to decouple the
+   multidimensional problems into a sequence of one-dimensional problems.
+   Sum-factorization enables fast operator evaluation, and the fast
+   diagonalization method addresses the efficient solution of problems
+   structured subdomains.  On unstructured subdomains, an optimal strategy
+   using static condensation with FDM bases was proposed in our previous
+   work.
+\item For simplices, Bernstein--Bezier expansions enable sum-factorized
+   operator evaluation, but there is no knowledge of basis transformation
+   that can diagonalize a problem on any structured or unstructured subdomain.
+\end{itemize}
+
+Other approaches
+\begin{itemize}
+\item Low-order-refined preconditioning, not fully understood/far from optimal.
+\item Sparse hierarchical bases from Beuchler \& Schoberl constructed from integrated Jacobi polynomials. 
+   These bases still couple interior DOFs, making static condensation difficult.
+\end{itemize}
+
+Our approach, main solver design ingredients:
+\begin{itemize}
+\item Construction of degrees of freedom that promote sparsity in the reference element, 
+   which in turn increases the efficiency of block preconditioners/decoupling strategies. The key idea is
+      to use the degrees of freedom proposed by Demkowicz with eigenbases for the space of bubbles on each subentity.
+\item Construction of space decompositions around small subdomains constructed around vertices and edges.
+\end{itemize}
+
 \section{Sparsity-promoting discretizations} \label{sec:dofs}

### 2024-06-18 paper/paper.tex
propagate notation

@@
 \item For simplices, Bernstein--Bezier expansions enable sum-factorized
-   operator evaluation, but there is no knowledge of basis transformation
+   operator evaluation, but there is no knowledge of a basis transformation
    that can diagonalize a problem on any structured or unstructured subdomain.
@@
 \begin{equation}
-X^k_p(\mesh) = \{ v \in H\Lambda^k(\Omega): \Forall K \in \mesh \Exists \hat{v} \in X^k_p(\Khat) \text{ s.t.} \left.v\right|_K = \pullback^k_K(\hat{v}) \},
+X^k_{h,p} = X^k_p(\mesh) = \{ v \in H\Lambda^k(\Omega): \Forall K \in \mesh \Exists \hat{v} \in X^k_p(\Khat) \text{ s.t.} \left.v\right|_K = \pullback^k_K(\hat{v}) \},
 \end{equation}
@@
 \begin{equation}
-V^1_{p, p_E}(K) = \{v\in [\P_p(K)]^d : v\cdot\bt \in \P_{p_E}(E) \}
+   X^1_{p, p_E}(K) = \{v\in [\P_p(K)]^d : \tr v \in \P_{p_E}(E) \Forall E \in \text{edges}(K)  \}
 \end{equation}
@@
 
-We define a basis for the dual of $V^1_{p, p_E}(K)$ as 
+We define a basis for the dual of $X^1_{p, p_E}(K)$ as 
 tangential moments along $E\in \text{edges}(\Khat)$,
@@
 where $\curl_S$ is the tangential curl on $S$,
-$\{\phi^S_j\}$ are the $H_0(\grad, S)$ bases constructed in \eqref{eq:hgrad-fdm},
-and $\{\curl_S\Phi^S_j\}$ is a basis for $\curl_S \mathbb{B}_{p_S}(\curl, S)$
+$\{\phi^S_j\}$ are the $B^0_{p_S}(S)$ bases constructed in \eqref{eq:hgrad-fdm},
+and $\{\curl_S\Phi^S_j\}$ is a basis for $\curl_S B^1_{p_S}(S)$
 such that
@@
    (\curl_S\Phi^S_j, \curl_S\Phi^S_i)_{S} = \delta_{ij}, \quad
-   (\Phi^S_j, \Phi^S_i)_{S} = \lambda_j\delta_{ij}, \quad
-   \Phi^S_j\times\bn = 0 \text{ on }\partial S.
+   (\Phi^S_j, \Phi^S_i)_{S} = \lambda_j\delta_{ij}.
 \end{equation}
@@
 \begin{equation}
-V^2_{p, p_F}(K) = \{v\in [\P_p(K)]^d : v\cdot\bn \in \P_{p_F}(F) \}
+   X^2_{p, p_F}(K) = \{v\in [\P_p(K)]^d : \tr v \in \P_{p_F}(F) \Forall F \in \text{faces}(K)  \}
 \end{equation}
@@
 
-We define a basis for the dual of $V^2_{p, p_F}(K)$ as normal moments on $F\in
+We define a basis for the dual of $X^2_{p, p_F}(K)$ as normal moments on $F\in
 \text{faces}(\Khat)$,
@@
 \end{align}
-where $\{\Phi^K_j\}$ are the $H_0(\curl, S)$ bases constructed in
-\eqref{eq:hcurl-fdm}, and $\{\div\Psi^K_j\}$ is a basis for $\div
-\mathbb{B}_p(\div, K)$ such that
+where $\{\Phi^K_j\}$ are the $B^1_p(K)$ bases constructed in
+\eqref{eq:hcurl-fdm}, and $\{\div\Psi^K_j\}$ is a basis for $\div B^2_p(K)$ such that
 \begin{equation} \label{eq:hdiv-fdm}
    (\div\Psi^K_j, \div\Psi^K_i)_K = \delta_{ij}, \quad
-   (\Psi^K_j, \Psi^K_i)_{K} = \lambda_j\delta_{ij}, \quad
-   \Psi^K_j\cdot\bn = 0 \text{ on }\partial K.
+   (\Psi^K_j, \Psi^K_i)_{K} = \lambda_j\delta_{ij}.
 \end{equation}
@@
 \begin{equation}
-   V^0_{h,p} = \left.V^0_{h,p}\right|_\mathcal{I} + V^0_{h,1} 
-   + \sum_{v\in E^0(\mesh)} \left.\tilde{V}^0_{h,p}\right|_{\star v} 
+   X^0_{h,p} = \left.X^0_{h,p}\right|_\mathcal{I} + X^0_{h,1} 
+   + \sum_{v\in E^0(\mesh)} \left.\tilde{X}^0_{h,p}\right|_{\star v} 
 \end{equation}
@@
 \begin{equation}
-   V^1_{h,p} = \left.V^1_{h,p}\right|_\mathcal{I} +  V^1_{h,1}
-   + \sum_{e\in E^1(\mesh)} \left.\tilde{V}^1_{h,p}\right|_{\star e} 
-   + \sum_{v\in E^0(\mesh)} \mathrm{d} \left.\tilde{V}^0_{h,p}\right|_{\star v} 
+   X^1_{h,p} = \left.X^1_{h,p}\right|_\mathcal{I} +  X^1_{h,1}
+   + \sum_{e\in E^1(\mesh)} \left.\tilde{X}^1_{h,p}\right|_{\star e} 
+   + \sum_{v\in E^0(\mesh)} \mathrm{d} \left.\tilde{X}^0_{h,p}\right|_{\star v} 
 \end{equation}
@@
 \begin{equation}
-   V^2_{h,p} = \left.V^2_{h,p}\right|_\mathcal{I} +  V^2_{h,1} 
-   + \sum_{e\in E^1(\mesh)} \left.\tilde{V}^2_{h,p}\right|_{\star e}. 
+   X^2_{h,p} = \left.X^2_{h,p}\right|_\mathcal{I} +  X^2_{h,1} 
+   + \sum_{e\in E^1(\mesh)} \left.\tilde{X}^2_{h,p}\right|_{\star e}. 
 \end{equation}

### 2024-06-19 paper/paper.tex
...

@@
 together the interior and interface degrees of freedom. Thus, on the
-reference simplex, the Schur complement resulting from elimination of
+reference simplex, the Schur complement resulting from the elimination of
 interior degrees of freedom is simply the interface block itself. This
@@
 \section{Introduction} \label{sec:introduction}
+
+High-order finite element discretizations have very attractive properties, such as rapid convergence and higher arithmetic intensity.
+However, their implementation requires great care, as na\"ive approaches to operator evaluation and linear system solution can scale badly in terms of memory and floating-point operations (flops) as the polynomial degree $p$ increases.
+On a single $d$-dimensional cell, the number of degrees of freedom scales like $p^d$; computing the $\mathcal{O}(p^{2d})$ entries of a dense stiffness matrix with standard Lagrange-type elements would cost $\mathcal{O}(p^{3d})$ flops.
+Making high-order discretizations competitive therefore requires better algorithms.
+
+For tensor-product cells (quadrilaterals and hexahedra), a tensor-product basis naturally exposes the required structure to decouple multidimensional problems into a sequence of one-dimensional problems, reducing computational cost and storage.
+For example, the sum-factorization assembly strategy enables fast matrix-free operator evaluation in $\mathcal{O}(p^{d+1})$ operations and $\mathcal{O}(p^{d})$ storage~\cite{orszag80}, and the fast diagonlization method (FDM) addresses the solution of separable problems on structured domains with the same complexities~\cite{lynch64}.
+The previous work of the first two authors employing FDM-inspired finite element bases with suitable space decompositions yields the solution of certain separable and non-separable problems on unstructured domains with the same complexities~\cite{brubeck22,brubeck23}.
+
+Achieving similar computational complexities on simplices is more challenging.
+Bernstein--B\'ezier expansions enable sum-factorized matrix-free operator evaluation in $\mathcal{O}(p^{d+1})$ operations and $\mathcal{O}(p^{d})$ storage~\cite{ainsworthXX}, but there is no known analogue of an FDM approach that solves the linear systems arising from separable problems with these complexities.
+This manuscript makes a step towards addressing this challenge.
+
+The essential idea of our approach is to devise new finite elements for the $L^2$ de Rham complex that promote sparsity in the reference
+element, in a similar spirit to~\cite{brubeck23}.
+On the reference cell, the interior and interface degrees of freedom are decoupled; on generally mapped cells, they are only coupled by the geometric transformation, and hence for suitably-shaped cells the coupling is weak.
+This increases the efficiency of block preconditioning strategies that decouple the interface and interior degrees of freedom.
+This reference cell sparsity is achieved by employing the framework of Demkowicz et al.~\cite{demkowicz00}, by making a clever choice for bases of bubble spaces that arise in the construction of the degrees of freedom.
+The bases for the bubble spaces are constructed by solving a handful of offline eigenproblems on the reference cell for each polynomial degree $p$.
+This combination of finite elements and block preconditioning strategy solves the Riesz maps of the $L^2$ de Rham complex in $\mathcal{O}(p^{3(d-1)})$ operations and $\mathcal{O}(p^{2(d-1)})$ storage; worse than the optimal complexities achieved on tensor-product cells, but orders of magnitude better than previously available approaches [FIXME is this true].
 

### 2024-06-27 paper/paper.tex
Fix references

@@
 For example, the sum-factorization assembly strategy enables fast matrix-free operator evaluation in $\mathcal{O}(p^{d+1})$ operations and $\mathcal{O}(p^{d})$ storage~\cite{orszag80}, and the fast diagonlization method (FDM) addresses the solution of separable problems on structured domains with the same complexities~\cite{lynch64}.
-The previous work of the first two authors employing FDM-inspired finite element bases with suitable space decompositions yields the solution of certain separable and non-separable problems on unstructured domains with the same complexities~\cite{brubeck22,brubeck23}.
+The previous work of the first two authors employing FDM-inspired finite element bases with suitable space decompositions yields the solution of certain separable and non-separable problems on unstructured domains with the same complexities~\cite{brubeck22,brubeck24}.
 
 Achieving similar computational complexities on simplices is more challenging.
-Bernstein--B\'ezier expansions enable sum-factorized matrix-free operator evaluation in $\mathcal{O}(p^{d+1})$ operations and $\mathcal{O}(p^{d})$ storage~\cite{ainsworthXX}, but there is no known analogue of an FDM approach that solves the linear systems arising from separable problems with these complexities.
+Bernstein--B\'ezier expansions enable sum-factorized matrix-free operator evaluation in $\mathcal{O}(p^{d+1})$ operations and $\mathcal{O}(p^{d})$ storage~\cite{ainsworth11}, but there is no known analogue of an FDM approach that solves the linear systems arising from separable problems with these complexities.
 This manuscript makes a step towards addressing this challenge.
@@
 The essential idea of our approach is to devise new finite elements for the $L^2$ de Rham complex that promote sparsity in the reference
-element, in a similar spirit to~\cite{brubeck23}.
+element, in a similar spirit to~\cite{brubeck24}.
 On the reference cell, the interior and interface degrees of freedom are decoupled; on generally mapped cells, they are only coupled by the geometric transformation, and hence for suitably-shaped cells the coupling is weak.

### 2024-06-27 paper/paper.tex
Solver diagram

@@
 
-\section{Space decompositions}
+\section{Design of preconditioners}
 
@@
 
+\begin{figure}
+\footnotesize
+\centering
+\begin{tikzpicture}[%
+ every node/.style={draw=black, thick, anchor=west},
+grow via three points={one child at (-0.0,-0.7) and
+	two children at (0.0,-0.7) and (0.0,-1.4)},
+edge from parent path={(\tikzparentnode.210) |- (\tikzchildnode.west)}]
+\node {KSP: Conjugate Gradients}
+   child {node {PC: Two-level vertex-star additive Schwarz}
+   child {node {$p$-fine: Interior-facet additive decomposition}
+      child {node {Interior: point-Jacobi}}
+      child {node {Facet: vertex-star patches}}
+   }
+   child[missing]{}
+   child[missing]{}
+   child {node {$p$-coarse ($p=1$): Cholesky}
+      %child {node {Relaxation: Hiptmair--Jacobi}}
+      %child {node {$h$-coarse: Cholesky}}
+   }
+};
+\end{tikzpicture}
+\caption{Solver diagram for $\Hgrad$ using the interior-facet decomposition.}
+\end{figure}
+
 \section{Numerical results} \label{sec:results}

### 2024-06-27 paper/paper.tex
Sketch section 3

@@
 \section{Design of preconditioners}
+In this section we desribe the solver components and their computational complexities.
 
+\begin{itemize}
+\item The bases described above present weak coupling between the interior and facet degrees of freedom.
+\item We propose a block diagonal preconditioner that only assembles the facet block, and only requires the diagonal entries of the interior block.
+\item The Krylov solver only requires matrix-free operation evaluation and preconditioner application.
+It is not evident that these bases expose structure for sum-factorization as the Bernstein--Bezier bases, hence operator application requires $\bigo{p^{2d}}$ flops and $\bigo{p^{2d}}$ storage.
+The storage cost here comes from tabulating the basis and its derivatives onto the quadrature points on the reference cell and does not scale with the number of cells.
+\item The facet cell submatrices have $\bigo{p^{2(d-1)}}$ entries, and cost $\bigo{p^{3(d-1)}}$ flops to assemble. 
+The solution of a patch problem on the facets involves the factorization of the patch matrix, which requires $\bigo{p^{2(d-1)}}$ storage and $\bigo{p^{3(d-1)}}$ flops top compute, solving the patch subproblems at each Krylov iteration incurs $\bigo{p^{2(d-1)}}$ flops per patch.
+\item In three dimensions the costs of operator application and preconditioner setup and application balance to $\bigo{p^6}$ flops and $\bigo{p^4}$ storage.
+\end{itemize}
+
+\subsection{Choice of space decompositions}
 We follow 
@@
 
-\begin{figure}
+\begin{figure}[htbp]
 \footnotesize

### 2024-06-27 paper/paper.tex
Writing with Pablo

@@
 \newcommand{\pef}[1]{\todo[inline,color=blue!40]{PEF: #1}}
+\newcommand{\pbm}[1]{\todo[inline,color=red!40]{PEF: #1}}
 
@@
 
-We discretize the $\Ltwo$ de Rham complex on triangular and tetrahedral meshes with finite element subcomplexes of the first and second kinds.
+We discretize the $\Ltwo$ de Rham complex on triangular and tetrahedral meshes with finite element subcomplexes of the first and second kinds (\Cref{fig:3DL2deRham}).
 
@@
 \end{tikzcd}
-\caption{The $\Ltwo$ de Rham complex (middle), and the finite element subcomplexes of the first (above) and second kinds (below).}
+\caption{The $\Ltwo$ de Rham complex in three dimensions (middle), and the finite element subcomplexes of the first (above) and second kinds (below).}
+\label{fig:3DL2deRham}
 \end{figure}
@@
 
-For each finite element subcomplex built on a mesh $\mesh$ of a domain $\Omega \subset \mathbb{R}^d$, we denote the spaces discretizing $k$-forms $H\Lambda^k(\Omega)$ with polynomial degree $p$ as $X^k_p(\mesh)$.
-These are built from cellwise spaces via the standard pullback construction from a reference cell $\Khat$:
+For each finite element subcomplex built on a mesh $\mesh$ of a bounded Lipschitz domain $\Omega \subset \mathbb{R}^d$, we denote the spaces discretizing $k$-forms $H\Lambda^k(\Omega)$ with polynomial degree $p$ as $X^k_p(\mesh)$.
+We employ the same notation for spaces of the first or second kind where the distinction does not matter.
+These spaces are built from cellwise spaces $X^k_p(\Khat)$ (constructed below) via the standard pullback construction from a reference cell $\Khat$:
 \begin{equation}
@@
 \begin{equation}
-B^k_p(S) \coloneqq \{v \in X^k_p(S): \tr v = 0 \text{ on } \partial S\}.
+B^k_p(S) \coloneqq \{\left.v\right|_S \in X^k_p(\Khat): \tr \left(\left.v\right|_S\right) = 0 \text{ on } \partial S\}.
 \end{equation}
 
-Throughout we choose $\Khat$ to be an equilateral simplex, as this ...
+Throughout we choose $\Khat$ to be an equilateral simplex, as this simplifies the construction of degrees of freedom on facet entities.
+We denote the entities in $\Khat$ of dimension $k$ by $\Delta^k(\Khat)$.
 
@@
 We define a basis for the dual of $X^0_p(\Khat)$ as
-point evaluations at $V \in \text{vertices}(\Khat)$,
+point evaluations at vertices $V \in \Delta^0(\Khat)$,
 \begin{equation}
@@
 \end{equation}
-and for each sub-entity $S \in \text{edges}(\Khat) \union \text{faces}(\Khat) \union \text{interior}(\Khat)$
+and for each higher-dimensional sub-entity $S \in \cup_{k=1}^d \Delta^k(\Khat)$
 we take integral moments of surface gradients, 
@@
 \end{equation}
-The eigenbases $\{\phi^S_j\}$ are numerically computed offline and only once 
-on the reference interval, triangle, and tetrahedron.
+The eigenbasis $\{\phi^S_j\}$ is numerically computed offline and only once
+for each kind of reference subentity of dimension greater than or equal to one. Since $\Khat$ is symmetric, only one facet, edge, etc.~needs to be computed, and the rest are constructed by congruence.
 
@@
 Recall the definition of the $\Hcurl$-conforming finite element spaces
+\pef{Find this in a book somewhere and cite it}
 \begin{equation}
-   X^1_{p, p_E}(K) = \{v\in [\P_p(K)]^d : \tr v \in \P_{p_E}(E) \Forall E \in \text{edges}(K)  \}
+   X^1_{p}(\Khat) = \{v\in [\P_p(K)]^d : \tr v \in \P_{p_E}(E) \Forall E \in \Delta^1(\Khat)  \}.
 \end{equation}
-for $p_E=p-1$ we obtain the \Nedelec space of the first kind $\Ned_p(K)$, while
-$p_E=p$ gives the \Nedelec space of the second kind $\NedTwo_p(K) = [\P_p(K)]^d$. 
+Choosing $p_E=p-1$ we obtain the \Nedelec space of the first kind $\Ned_p(\Khat)$, while
+choosing $p_E=p$ gives the \Nedelec space of the second kind $\NedTwo_p(K) = [\P_p(\Khat)]^d$.
+\pef{Cite \Nedelec}
 These spaces enforce continuity of the tangential trace across mesh interfaces.
 
-We define a basis for the dual of $X^1_{p, p_E}(K)$ as 
-tangential moments along $E\in \text{edges}(\Khat)$,
+We define a basis for the dual of $X^1_p(K)$ as 
+tangential moments along edges $E\in \Delta^1(\Khat)$,
 \begin{equation}
@@
 \end{equation}
-and for each sub-entity $S \in \text{faces}(\Khat) \cup \text{interior}(\Khat)$,
+and for each higher-dimensional sub-entity $S \in \cup_{k=2}^d \Delta^k(\Khat)$,
 \begin{align}
@@
 where $\curl_S$ is the tangential curl on $S$,
-$\{\phi^S_j\}$ are the $B^0_{p_S}(S)$ bases constructed in \eqref{eq:hgrad-fdm},
+$\{\phi^S_j\}$ is the basis for $B^0_{p_S}(S)$ constructed in \eqref{eq:hgrad-fdm},
 and $\{\curl_S\Phi^S_j\}$ is a basis for $\curl_S B^1_{p_S}(S)$
-such that
+constructed by solving the following problem:
+find $\{\Phi^S_j\} \subset B^1_{p_S}(S)$ such that
 \begin{equation} \label{eq:hcurl-fdm}
    (\curl_S\Phi^S_j, \curl_S\Phi^S_i)_{S} = \delta_{ij}, \quad
-   (\Phi^S_j, \Phi^S_i)_{S} = \lambda_j\delta_{ij}.
+   (\Phi^S_j, \Phi^S_i)_{S} = \lambda_j\delta_{ij}, \quad \lambda_j > 0.
 \end{equation}
+\pef{FIXME: what is $p_S$?}
+
+The eigenbasis $\{\Phi^S_j\}$ is numerically computed offline and only once for each kind of reference subentity of dimension greater than or equal to two.
 
